---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggsurvfit

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ggsurvfit)](https://CRAN.R-project.org/package=ggsurvfit)
[![Codecov test coverage](https://codecov.io/gh/ddsjoberg/ggsurvfit/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ddsjoberg/ggsurvfit?branch=main)
[![R-CMD-check](https://github.com/ddsjoberg/ggsurvfit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ddsjoberg/ggsurvfit/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Introduction

The {ggsurvfit} package eases the creation of time-to-event (aka survival) endpoint figures with ggplot.
The concise and modular code creates images that are ready for sharing or publication.
Each of the functions that add to or modify the existing ggplot are written as proper ggplot geoms or stat methods, allowing the functions from this package to be combined with functions from ggplot and other ggplot extensions.

It's recommended to use the `survfit2()` function with this package, rather than `survival::survfit()` to ensure all quantities are able to be calculated from any environment the functions are called.

## Installation

You can install the development version of ggsurvfit from [GitHub](https://github.com/ddsjoberg/ggsurvfit) with:

``` r
# install.packages("devtools")
devtools::install_github("ddsjoberg/ggsurvfit")
```

## Examples

Let's begin with an example illustrating a common Kaplan-Meier survival curve.

```{r example, fig.height=4.5}
library(ggsurvfit)
library(ggplot2)

p <- 
  survfit2(Surv(time, status) ~ sex, data = df_lung) |>
  ggsurvfit() +
  add_censor_mark() +
  add_confidence_interval() +
  add_risktable() +
  add_quantile(color = "gray40")
p
```

<hr>
Because each of the functions are written as proper ggplot geoms, you can add any ggplot function to modify the figure.

```{r example-styled}
p +
  theme_bw() + 
  theme(legend.position = "bottom") +
  labs(
    y = "Probability of survival",
    x = "Months since treatment",
    title = "Kaplan-Meier Estimate of Survival by Sex",
    # remove the fill and color legend labels (Sex appears in title)
    fill = "", color = ""
  ) + 
  # reduce padding on edges of figure, and format axes
  scale_y_continuous(label = scales::percent, expand = c(0.01, 0)) +
  scale_x_continuous(breaks = 0:5*6, expand = c(0.02, 0))
```

<hr>
The package also supports unstratified models and you can add multiple quantile guidelines.

```{r example-unstrat, fig.height=3.5}
survfit2(Surv(time, status) ~ 1, data = df_lung) |>
  ggsurvfit() +
  add_quantile(linetype = 3) +
  add_quantile(y_value = 0.25, linetype = 2) +
  add_confidence_interval() +
  add_risktable()
```

<hr>
You can even facet...but you _cannot_ facet with the risktable, however.

```{r example-facet, fig.height=3.5}
survfit2(Surv(time, status) ~ sex, data = df_lung) |>
  ggsurvfit() +
  add_censor_mark(shape = 4) +
  add_quantile(linetype = 3) +
  add_confidence_interval() +
  ggplot2::facet_grid(~strata)
```

<hr>
The package works seamlessly with other ggplot extension package, such as {gghighlight}.

```{r example-gghighlight, fig.height=3.5}
survfit2(Surv(time, status) ~ ph.ecog, data = df_lung) |> 
  ggsurvfit() +
  ggplot2::labs(color = "Gender") +
  gghighlight::gghighlight(strata == "Asymptomatic", calculate_per_facet = TRUE)
```

<hr>
The package also plots cumulative incidence estimates in the presence of competing events.

```{r example-cuminc}
library(tidycmprsk)

cuminc(Surv(ttdeath, death_cr) ~ trt, trial) %>%
  ggcuminc(outcome = "death from cancer") +
  add_confidence_interval() +
  add_quantile(y_value = 0.20) +
  add_risktable() +
  theme_bw() + 
  labs(x = "Months Since Treatment") +
  theme(legend.position = "bottom") +
  scale_y_continuous(label = scales::percent, expand = c(0.02, 0)) +
  scale_x_continuous(breaks = 0:4 * 6, expand = c(0.02, 0))
```

```{r, echo=FALSE}
knitr::knit_exit()
```

## Related Packages

```{r, echo=FALSE, size="tiny"}
dplyr::tribble(
  ~Package,             ~CRAN, ~`Risk Table`,
  "survminer",          "X",   "X",
  "`galley::ggsurv()`", "X",   "",
  "ggfortify",          "X",   "",
  "visR",               "X",   "X",
  "iwillsurvive",       "",    "X",
  "KMunicate",          "X",   "X"
) %>%
  knitr::kable()
```
