---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE
)
```

# ggsurvfit

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/ddsjoberg/ggsurvfit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ddsjoberg/ggsurvfit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ddsjoberg/ggsurvfit/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ddsjoberg/ggsurvfit?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/ggsurvfit)](https://CRAN.R-project.org/package=ggsurvfit)
<!-- badges: end -->

## Introduction

The **ggsurvfit** package eases the creation of time-to-event (aka survival) endpoint figures with ggplot2.
The concise and modular code creates images that are ready for publication or sharing.
Competing risks cumulative incidence is also supported via `ggcuminc()`.

## Why ggsurvfit?

- **Use ggplot2 functions:** Each **ggsurvfit** add-on function (e.g. `add_confidence_interval()`, `add_risktable()`, etc.) is written as a proper ggplot2 ['geom'](https://ggplot2.tidyverse.org/reference/index.html), meaning that the package functions can be woven with ggplot2 functions seamlessly. You don't need to learn how to style the plot within the ggsurvfit functions: you may rely on the suite ggplot2 functions you have already learned.

- **Publishable Legends:** Raw variable names do not appear in the figure legend, e.g. `"sex=Female"`.

- **Limitless Customization:** You can modify the x-axis scales or any other plot feature and the risk table will still align with the plot.

## Installation

You can install the development version of **ggsurvfit** from [GitHub](https://github.com/ddsjoberg/ggsurvfit) with:

``` r
# install.packages("devtools")
devtools::install_github("ddsjoberg/ggsurvfit")
```

## Examples

Review the [**figure gallery**](http://www.danieldsjoberg.com/ggsurvfit/articles/gallery.html) for many more examples. 


```{r example, fig.height=5.5}
library(ggsurvfit)
library(ggplot2)

survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  # build Kaplan-Meier plot ----------------------------------------------------
  ggsurvfit(size = 1) +
  add_confidence_interval() +
  add_risktable() +
  add_quantile(y_value = 0.6, color = "gray50", size = 0.75) +
  
  # use ggplot2 functions to style the plot and update the labels --------------
  # limit plot to show 8 years and less
  coord_cartesian(xlim = c(0, 8)) +
  # update figure labels/titles
  labs(
    y = "Percentage survival",
    title = "Recurrence by Time Since Surgery to Randomization",
  ) +
  # reduce padding on edges of figure, and format axes
  scale_y_continuous(label = scales::percent, 
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0.015, 0)) +
  scale_x_continuous(breaks = 0:10, 
                     expand = c(0.02, 0))
```

## `survfit2()` vs `survfit()`

Both functions have identical inputs, so why do we need `survfit2()`?
The `survfit2()` tracks the environment from which the function was called, resulting in the following benefits.

  - We can reliably remove the raw variable names from the figure legend, e.g. `SEX=Female`.
  - P-values can be calculated with `survfit_p()` and added to figures.
  - The items above are often *possible* using `survfit()`. However, by utilizing the calling [environment](https://adv-r.hadley.nz/environments.html) we are assured the correct elements are found, rather than crossing our fingers that the search path contains the needed elements.

## CDISC ADaM ADTTE

The package also includes a gem for those using the [CDISC ADaM ADTTE](https://www.cdisc.org/standards/foundational/adam/adam-basic-data-structure-bds-time-event-tte-analyses-v1-0) data model.
The event indicator in ADTTE data sets is named `"CNSR"` and is coded in the opposite way the survival package expects outcomes---`1 = 'censored'` and `0 = 'event'`.
This difference creates an opportunity for errors to be introduced in an analysis.
The package exports a function called `Surv_CNSR()` to resolve this concern.

The function creates a survival object (e.g. `survival::Surv()`) that uses CDISC ADaM ADTTE coding conventions and converts the arguments to the status/event variable convention used in the survival package for both the event indicator and the time component.
The `"AVAL"` and `"CNSR"` arguments are passed to `survival::Surv(time = AVAL, event = 1 - CNSR, type = "right", origin = 0)`.
The function can be used in **ggsurvfit** as well as any other package that uses `survival::Surv()`.

```{r}
survfit(Surv_CNSR() ~ 1, adtte)
```

```{r, echo=FALSE}
knitr::knit_exit()
```

## Related Packages

```{r, echo=FALSE, size="tiny"}
dplyr::tribble(
  ~Package,             ~CRAN, ~`Risk Table`,
  "survminer",          "X",   "X",
  "`galley::ggsurv()`", "X",   "",
  "ggfortify",          "X",   "",
  "visR",               "X",   "X",
  "iwillsurvive",       "",    "X",
  "KMunicate",          "X",   "X"
) %>%
  knitr::kable()
```
