---
title: "Gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The gallery exhibits the default plots as well as many modifications one can make.

```{r setup}
library(ggsurvfit)
library(ggplot2)
library(patchwork)

km_estimate <- survfit2(Surv(AVAL, 1 - CNSR) ~ STR01, data = adtte)
```

## Modifications with ggplot2

Let's begin with showing the default plot and common modifications that are made with ggplot2 functions.
- Add the percent sign to the y-axis label
- Reduce padding in the plot area around the curves
- Add additional tick marks on the x-axis
- Update color of the lines
- Using the ggplot2 minimal theme
- Move legend to the top of plot

```{r}
gg_default <-
  km_estimate %>%
  ggsurvfit() +
  add_confidence_interval() +
  labs(title = "Default")

gg_styled <-
  gg_default +
  scale_y_continuous(labels = scales::percent, expand = c(0.01, 0)) +
  scale_x_continuous(breaks = 0:5, expand = c(0.02, 0)) +
  scale_color_manual(values = c('#54738E', '#82AC7C')) +
  scale_fill_manual(values = c('#54738E', '#82AC7C')) +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(title = "Modified")

gg_default + gg_styled
```

## Grey-scale Figures

Commonly, you need a black and white figure, and that is achieved using grey-scale ggplot2 functions.

```{r}
km_estimate %>%
  ggsurvfit() +
  add_confidence_interval() +
  scale_color_grey() +
  scale_fill_grey() +
  labs(title = "Grey Scale")
```

## KMunicate

To get figures that align with the guidelines outlined in ["Proposals on Kaplan–Meier plots in medical research and a survey of stakeholder views: KMunicate."](http://dx.doi.org/10.1136/bmjopen-2019-030215), use the `theme_ggsurvfit_KMunicate()` theme along with these function options.

```{r}
km_estimate %>%
  ggsurvfit(linetype_aes = TRUE) +
  add_confidence_interval() +
  add_risktable(
    risktable_stats = c("n.risk", "cum.censor", "cum.event"),
    risktable_height = 0.2
  ) +
  theme_ggsurvfit_KMunicate() +
  scale_x_continuous(breaks = 0:5, expand = c(0.02, 0)) +
  theme(legend.position = c(0.85, 0.85))
```

## Risk Tables

The default risk table styling is ready for publication.

```{r}
km_estimate %>%
  ggsurvfit() +
  add_confidence_interval() +
  add_risktable()
```

You can also group the risktable by the statistics rather than the stratum.

```{r}
km_estimate %>%
  ggsurvfit() +
  add_confidence_interval() +
  add_risktable(risktable_group = "risktable_stats")
```

Use `add_risktable_strata_symbol()` to replace long stratum labels with a color symbol.

Without `add_risktable_strata_symbol()`:

```{r}
km_long_labels <- 
  survfit2(Surv(AVAL, 1 - CNSR) ~ TRT01P, data = adtte) %>%
  ggsurvfit() +
  add_risktable(
    risktable_stats = "n.risk", 
    risktable_group = "risktable_stats",
    risktable_height = 0.2
  ) +
  theme(legend.position = "right") +
  labs(color = NULL)
km_long_labels
```

With `add_risktable_strata_symbol()`:

```{r}
km_long_labels +
  add_risktable_strata_symbol()
```

## Quantiles and Censor Markings

Add guidelines for survival quantiles and markings for censored patients using `add_quantile()` and `add_censor_mark()`.

```{r}
km_estimate %>%
  ggsurvfit(size = 0.8) +
  add_censor_mark(size = 2, alpha = 0.2) +
  add_quantile(y_value = 0.5, linetype = "dotted", color = "grey50", size = 0.8) +
  add_quantile(y_value = 0.75,  color = "grey50", size = 0.8) 
```

## Transformations

Show the probability of an event rather than the probability of being free from the event with transformations.
Custom transformations are also available.

```{r}
km_estimate %>%
  ggsurvfit(type = "risk", size = 0.8) +
  add_confidence_interval()
```

## Faceting

Curves created with ggsurvfit can also later be faceted using ggplot2.
Note, however, that faceted curves cannot include a risk table.

```{r}
km_estimate %>%
  ggsurvfit() +
  add_confidence_interval() +
  facet_wrap(~strata, nrow = 1) +
  theme(legend.position = "none") +
  scale_x_continuous(n.breaks = 6) +
  labs(title = "Survival by Hormone Receptor Status")
```

