---
title: "Themes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Plot Themes

Both `ggsurvfit()` and `ggcuminc()` use `theme_ggsurvfit_default()` as the default ggplot2 theme.
The theme is similar to `ggplot2::theme_bw()`.
While this general theme is provides a nice-looking figure, you may want to modify it.

You'll use the `ggsurvfit(theme=)` argument to change the default theme.
You can pass any ggplot2 theme to this argument, or simply add a theme using typical ggplot2 syntax.

```{r}
library(ggsurvfit)
library(ggplot2)
library(patchwork)

km_estimate_sex <- survfit2(Surv(time, status) ~ sex, data = df_lung)

gg_theme_default1 <- 
  km_estimate_sex %>%
  ggsurvfit(theme = theme_ggsurvfit_default()) +
  labs(title = "Using `theme=` argument")

gg_theme_default2 <- 
  km_estimate_sex %>%
  ggsurvfit(theme = NULL) +
  theme_ggsurvfit_default() +
  labs(title = "Using ggplot `+`")

gg_theme_default1 + gg_theme_default2
```

As you can see, either method of passing the theme results in an identical figure.
You can easily apply any user-created theme or any theme that ships with the ggplot2 package.
To construct a custom theme, include ggplot2 calls in a list as in the example below.

```{r}
km_estimate_sex %>%
  ggsurvfit(theme = list(theme_classic(), theme(legend.position = "bottom"))) +
  labs(title = "Custom Plot Theme")
```

### Risk Table Themes

Similar to the plots above, the risk tables also come with a default theme.
To understand how to use risk table themes, one must first understand what a risk table is.
Each risk table that is placed below a plot is itself a ggplot created with `ggplot2::geom_text()`.

Unlike the plot themes, the risk table themes can only be passed via `add_risktable(theme=)`.

```{r}
km_estimate_sex %>%
  ggsurvfit() +
  add_risktable(
    risktable_stats = "n.risk",
    theme = theme_risktable_default()
  )
```

A typical theme appropriate for a risk table plot will remove gridlines and provide a clean area to display the numbers at risk.
In the risk table above, the strata levels are y-axis labels, and "At Risk" is the plot title.

Another risk table theme adds a box around the statistics presented.

```{r}
km_estimate_sex %>%
  ggsurvfit() +
  add_risktable(
    risktable_stats = "n.risk",
    theme = theme_risktable_boxed()
  )
````
It's possible to replace the stratum levels with a symbol, which is particularly helpful when you have long group labels.

```{r}
km_estimate_sex %>%
  ggsurvfit() +
  add_risktable(risktable_stats = "n.risk") +
  add_risktable_strata_symbol()
````

When customizing the risk table themes and using `add_risktable_strata_symbol()`, note that the symbol is added with `ggplot2::scale_y_discrete()`; including a theme that uses the discreet y scale could cause a conflict.

If you need a risk table theme that is not currently available in ggsurvfit, community-based themes are welcome!
Consider adding your theme to the package.


